---
title: "p8105_hw3_km3529"
author: Karina Myers
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(patchwork)
knitr::opts_chunk$set(
	fig.width = 6, 
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```



## Problem 1 
```{r}
data("instacart")
```
This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. 

Observations are the level of items in orders by user. There are user and order variables such as user ID, order ID, order day, and order hour. There are also item variables such as product name, aisle, department, whether the item has been ordered before, and some numeric codes. 


How many aisles are there, and which aisles are the most items ordered from?
```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```


Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

NOTE: cat var will be ordered numerically --> change to factor 

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle), 
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart %>% 
	filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4 ) %>% 
  arrange(aisle, rank) %>% 
knitr::kable()
```



Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour
  )
```





## Problem 2



This problem uses five weeks of accelerometer data collected on a 63 year-old male with BMI 25, who was admitted to the Advanced Cardiac Care Center of Columbia University Medical Center and diagnosed with congestive heart failure (CHF). The data can be downloaded here. In this spreadsheet, variables activity.* are the activity counts for each minute of a 24-hour day starting at midnight.


Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

```{r}
accel_df = 
  read_csv("./Data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
      activity_1:activity_1440, 
      names_to = "minute",
      names_prefix = "activity_", 
      values_to = "activity_count") %>% 
  mutate(
    day_type = case_when(
        day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday") ~ "weekday", 
        day %in% c("Saturday", "Sunday") ~ "weekend", 
        TRUE ~ "") , 
    activity_count = as.numeric(activity_count), 
    day = as.factor(day), 
    minute = as.numeric(minute), 
    day = forcats::fct_relevel(
        day, c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
    )
```

SUMMARY HERE 
  - During observation periods, the devices measure “activity counts” in a short period; one-minute intervals   are common. Because accelerometers can be worn comfortably and unobtrusively, they produce                    around-the-clock observations.
  - 5 weeks of data 


Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?
```{r}
accel_df %>% 
  group_by(day, week) %>% 
  summarize(daily_activity = sum(activity_count)) %>% 
  pivot_wider(
      names_from = day, 
      values_from = daily_activity) %>% 
  knitr::kable(digits = 1)
```

  - TRENDS????
  - weeks 2 and three hav e alot of activity? 


Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.
```{r}
accel_df %>% 
  ggplot(aes(x = minute, y = activity_count, color = day)) + 
 geom_line(alpha = .5) 

#add title to Plot
#notes: geom smooth?
# should have 35 lines? 
## get thrity five lines
```



## Problem 3 
This problem uses the NY NOAA data. DO NOT include this dataset in your local data directory; instead, load the data from the p8105.datasets package using:
```{r}
library(p8105.datasets)
data("ny_noaa")
```
 
WRITE DESCRIPTION

  - The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue. 
  - Then, do or answer the following (commenting on the results of each):

Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. 
```{r}
ny_noaa = 
  ny_noaa %>% 
  separate(date, into = c("year", "month", "day"), convert = TRUE) %>% 
  mutate(
    year = factor(year), 
    tmax = as.integer(tmax), 
    tmin = as.integer(tmin), 
    prcp = prcp/10, 
    tmax = tmax/10, 
    tmin = tmin/10
  )
```
prcp is measured in mm (originally measured in tenths of mm) 
tmin/tmax are in degrees celcius (originally measured in tenth of degree) 

For snowfall, what are the most commonly observed values? Why?
```{r}
ny_noaa %>% 
  drop_na(snwd) %>% 
  count(snwd) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 10)
```


Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?
```{r}
ny_noaa %>% 
  filter(month %in% c("1", "7")) %>% 
  group_by(month, year, id) %>% 
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = mean_tmax, group = id)) + 
    geom_point(size = .5) +
    geom_path(alpha = .5, size = .1) + 
    facet_grid(.~month) + 
    theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust = 1))


```
can you see global warming? 
are some stations always colder? 


Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.
```{r, fig.height=5, fig.width=10}
#Plot 1
tmin_tmax_p = 
  ny_noaa %>% 
    drop_na(tmin, tmax) %>% 
    ggplot(aes(x = tmax, y = tmin)) + 
    geom_hex() + 
    labs(
      title = "Maximum Temperature v Minimum Temperature", 
      x = "Maximum Temperature Per Day", 
      y = "Minimum Temperature Per Day" )

#plot 2
snow_p = 
  ny_noaa %>% 
  drop_na(snow) %>% 
  filter(snow < 100) %>% 
  filter(snow > 0 ) %>% 
    ggplot(aes(x = year, y = snow)) + 
    geom_violin(alpha = .5) + 
    theme(axis.text.x = element_text(angle = 270, vjust = 0.5, hjust = 1)) +   labs(
      title = "Snowfall by Year"
    )

tmin_tmax_p + snow_p

```











