---
title: "p8105_hw3_km3529"
author: Karina Myers
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse) 
library(p8105.datasets)

knitr::opt_chunk$set(
  fig.width = 6, 
  fig.asp = .6, 
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis", 
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 0 

```{r}
data("instacart") 
```

This datset contains  `r nrow(instacart)` rows and xolumns. 
observations are at the level of items in orders by users 


How many aisles, ane which are most items from? 
essentially a counting question 
```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

134 aisles 
1067 items come from air fresheners and candles 
ararnge lets you see where most are coming form 



let's a make a plot 
```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point()
```
hard to read soo 

deal with x axis so rotate x axis labels
```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

now let's put in an order that makes sense. 
ggplot put it from alphabetical. 
  - least common to most common 
```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle), 
    aisle = fct_reorder(aisle, n) 
  ) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


let's make a table
```{r}
instacart %>% 
  filter(aisle %in%  c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```

group by goes before count
the mutate line
  - using min_rank on n puts thigns in reverse order (smallest to largest)
  - descending does reverse order
  - 

%in% is this in this? 







pinkl ady apple and coffee ice cream 

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour
  )
```



## Problem 2

not tidy 
add variable 
```{r}
read_csv("./")
```

group by and summarize. then  make a table you would summarize 






